
\chapter{VEP-based brain-computer interface}

The previous chapter discussed the biology of the brain and described the brain potential called \gls{SSVEP}. The aim of this chapter is to describe which kind of visual stimuli can be used to elicit \gls{SSVEP} and how to detect \gls{SSVEP} in a \gls{EEG} recording. This knowledge is needed to extract information from the \gls{EEG} recording and use this information to control a robot. In other words, this chapter will discuss how to implement a \gls{VEP}-based \gls{BCI}.

\section{Visual stimulus}
\label{sec:stimuli}

As discussed in section~\ref{sec:VEP}, it is possible to present multiple visual stimuli to a subject and detect, which stimulus is the subject looking at. It was also mentioned that computer monitor can be used to present visual stimuli. This section will discuss how stimuli with certain blinking frequency can be displayed on a computer screen. Research has shown that LCD screens produce more reliable \gls{SSVEP} response than \gls{LED}~\cite{lcd_lcd_led}. Another article, however, concludes that \glspl{BCI} that used \glspl{LED} as visual stimuli have achieved better performance~\cite{ssvep_stim}. But since using \glspl{LED} requires dedicated hardware, in this thesis only the stimulus that can be presented by a computer monitor is discussed.

A computer monitor has certain size, resolution and \gls{refresh rate}. Monitor resolution and size limit the size of the visual stimuli that can be used and the distance between the stimuli. Monitor \gls{refresh rate}, on the other hand, limits the blinking frequency of the stimulus that can be used. Research has shown that LCD screens produce more reliable \gls{SSVEP} response when using monitor \gls{refresh rate} for measuring the time between stimuli presentations rather than a timer~\cite{lcd_lcd_led}. Therefore, monitor \gls{refresh rate} is used for timing also in this thesis.

Monitor \gls{refresh rate} is the number of consecutive images or \glspl{frame} shown on the screen in a second, assuming that the \glspl{frame} are produces at least as fast as they can be displayed. A \gls{frame} is one of the images that compose the changing picture on screen. Monitor with a \gls{refresh rate} of 60 Hz can display 60 \glspl{frame} per second. 

Often blinking one-coloured squares on one-coloured background are used as visual stimuli in \gls{SSVEP}-based \glspl{BCI}~\cite{ssvep_stim}. The squares may have symbols on them, for example letters or numbers. The stimuli of \gls{VEP}-based \gls{BCI} are also called \glspl{target}. The blinking of a target is called \gls{flickering}.

In every \gls{frame} each \gls{target} can be in one of two \glspl{state}---displayed or not displayed. The \gls{state} of a \gls{target} can be switched only when current \gls{frame} is replaced with next \gls{frame}. The \gls{state} switches should be distributed as evenly as possible for the target frequency to be constant. Distributing the \gls{state} switches is easier with some frequencies than others. For example, if \gls{refresh rate} is 60 Hz, then
\begin{itemize}
	\item 10 Hz \gls{flickering} can be achieved by presenting the \gls{target} $\frac{60}{10}=6$ times slower than the \gls{refresh rate}. This means, that the \gls{target} has to be presented once in every 6 \glspl{frame}. Since 6 is even number, 10 Hz \gls{flickering} can be achieved by changing the \gls{state} of the \gls{target} after every $\frac{6}{2} = 3$ \glspl{frame}. If this \gls{flickering} is plotted as a function of \gls{state} versus time as in figure~\ref{fig:flickering}, it can be seen that it produces a \gls{square wave}. In this thesis, the waveform produced by plotting the \gls{flickering} is called \gls{flickering waveform}.
	\item 12 Hz \gls{flickering} can be achieved by presenting the \gls{target} $\frac{60}{12}=5$ times slower than the \gls{refresh rate}. The \gls{target} has to be presented once in every 5 \glspl{frame} and since 5 is odd number, the amount of time the target is in on \gls{state} and in off \gls{state} cannot be equal. Therefore, the \gls{target} should be 3 \glspl{frame} in displayed \gls{state} and 2 \glspl{frame} in not displayed \gls{state} or the other way round. If representing this \gls{flickering} as a \gls{flickering waveform} as in figure~\ref{fig:flickering}, it can be seen that the waveform is a \gls{rectangular wave}.
	\item 11 Hz \gls{flickering} can be achieved by presenting the \gls{target} $\frac{60}{11}\approx 5.45$ times slower than the \gls{refresh rate}. This means, that the \gls{target} has to be presented once in every 5.45 \glspl{frame}. Since 5.45 is not a natural number, the \gls{flickering} will be irregular. 11 Hz target \gls{flickering} from the paper by Wang \textit{et al.}~\cite{11hz} is used as an example in figure~\ref{fig:flickering}. Although 11 Hz frequency produces irregular \gls{flickering}, it is still possible to detect \gls{SSVEP} elicited by it in the Emotiv EPOC recording~\cite{emotiv_11hz}.
\end{itemize}

\begin{figure}[h]
	\begin{subfigure}{\textwidth}
		\begin{tikztimingtable}[xscale=0.75, yscale=1.5, thick]
			10 Hz & [C] 10{3H 3L}\\
			11 Hz & [C] 11{2.73H 2.73L}\\
			12 Hz & [C] 12{2.5H 2.5L}\\
			\extracode
			\tablegrid[black!25,step=1]
		\end{tikztimingtable}
		\caption{Ideal target flickering}
	\end{subfigure}
	\begin{subfigure}{\textwidth}
		\begin{tikztimingtable}[xscale=0.75, yscale=1.5, thick]
			10 Hz & [C] 10{3H 3L}\\
			11 Hz & [C] 3H 3L 3H 2L 3H 3L 3H 2L 3H 3L 2H 3L 3H 3L 2H 3L 3H 3L 2H 3L 3H 2L\\
			12 Hz & [C] 12{3H 2L}\\
			\extracode
			\tablegrid[black!25,step=1]
		\end{tikztimingtable}
		\caption{Target flickering adjusted to refresh rate}
	\end{subfigure}
	\caption{Adjusting target flickering to refresh rate}
	\label{fig:flickering}
\end{figure}
A \gls{duty cycle} is used to characterise a \gls{rectangular wave}. \Gls{duty cycle} is the percentage of the amount of time the \gls{target} is in displayed \gls{state} in one period. If the target is in displayed \gls{state} for 2 \glspl{frame} and in not displayed \gls{state} for 3 \glspl{frame} in one period, then the \gls{duty cycle} of the \gls{rectangular wave} is $\frac{2}{2+3}\cdot 100\%=40\%$. \Gls{square wave} has a \gls{duty cycle} of 50\%. Research has shown that the \gls{SSVEP} elicited by \gls{square wave} \gls{flickering} can be more accurately detected than those elicited by \glspl{rectangular wave} \gls{flickering}~\cite{ssvep_stim}.

The previous discussion was about blinking shapes or \gls{single graphic} stimuli. There is another type of \gls{SSVEP} stimuli called \gls{pattern reversal} stimuli that can also be presented by computer screen. \Gls{pattern reversal} stimuli is rendered by changing between two different patterns, for example alternating the colours of a chequerboard~\cite{ssvep_stim}. The main difference between \gls{single graphic} stimuli and \gls{pattern reversal} stimuli is that \gls{single graphic} stimuli elicits \gls{SSVEP} response after every two alterations, while \gls{pattern reversal} stimuli elicits \gls{SSVEP} response after every alteration~\cite{ssvep_stim}.

The fastest possible target frequency can be achieved by changing between on and off \glspl{state} of a target every time a new \gls{frame} is displayed. If the \gls{state} is changed at lower rate, the \gls{target} frequency will also be lower. Perhaps it is most reasonable to calculate the \gls{target} frequency with:
\begin{equation}
	f_{single\mbox{ }graphic} = \frac{n}{2T} \qquad f_{pattern\mbox{ }reversal} = \frac{n}{T}
\end{equation}
where $T$ is the period of the \gls{flickering waveform}, $n$ is the number of times the \gls{target} \gls{state} is switched in a time period of $T$ and $f$ is the \gls{flickering} frequency.

Therefore computer monitor can be used to present visual stimuli to a subject and monitor \gls{refresh rate} should be used to measure the time between stimuli presentations. Adjusting \gls{target} \gls{flickering} to the \gls{frame} changes can produce different \glspl{flickering waveform} for different \gls{flickering} frequencies. Since the \glspl{flickering waveform} are different, the \gls{SSVEP} response will also be different.

\section{Fourier analysis}
\label{sec:fourier}

As discussed in the previous section, the \gls{flickering} of a \gls{target} can produce different waveforms. This section will discuss how these waveforms and other signals can be represented by sums of simpler trigonometric functions. The study of this decomposition process is called Fourier analysis, named after Joseph Fourier, whose insight to model all functions by trigonometric series was a breakthrough in the field in 1807.

The following paragraph is based on the book by Hartmann~\cite{pure_tone}. The simpler trigonometric functions that a signal is decomposed into are \glspl{pure tone}---the waveforms that contain only one frequency. All other waveforms have at least two frequencies. \Gls{pure tone} waveforms are sine and cosine waves. Important property of a \gls{pure tone} is that linear operations do not change the shape of the \gls{pure tone} waveform.

The \glspl{pure tone} are used to represent all possible frequencies that a signal may contain. The decomposition process of a signal is called \gls{Fourier transform} and it is used to decompose a function of time into \glspl{pure tone} or \glspl{frequency component} that make it up. The function of time can be for example the \gls{EEG} recording represented as voltage versus time or the \gls{target} \gls{flickering} represented as \gls{state} versus time. \Gls{Fourier transform} converts signal from time domain or the function of time to frequency domain or the function of frequency. The representation of a time-domain signal in a frequency domain is called \gls{frequency spectrum}. \Gls{frequency spectrum} contains information about amplitude and phase of different frequencies. Therefore, \gls{frequency spectrum} can be presented as a function of frequency versus amplitude and phase.

To represent both amplitude and phase, complex numbers are used. Complex numbers can be represented as a pair of real numbers and therefore complex numbers can be used to represent two values. The amplitude and phase do not correspond to the real and imaginary part of the complex number but rather are related to the absolute value or the modulus and phase of the complex number. In this thesis the phase information from the \gls{frequency spectrum} is not used. There are, however, \glspl{BCI} that also use the phase information~\cite{MPCC}. Since only the information about amplitude is required in this thesis, it is possible to convert the \gls{frequency spectrum} into \gls{power spectral density}, which is represented as a function of amplitude squared or power versus frequency.

To conclude previous discussion, the amount of frequency $f$ present in a signal can be calculated by first calculating the \gls{frequency spectrum} with \gls{Fourier transform} and then taking modulus squared of the \gls{frequency spectrum}'s value at frequency $f$.

In digital devices, however, theoretical \gls{power spectral density} cannot be calculated. The measurement period would have to be infinitely long to a acquire the true \gls{power spectral density}~\cite{psd}. Therefore, spectral estimation is used. The modulus squared of frequency spectrum in a real-world application is called \gls{periodogram} and it is the estimation of the \gls{power spectral density}. There are other spectral estimation methods available.

Since digital devices work with discrete signals as discussed in section~\ref{sec:EEG_comparison}, in a real-world application discrete version of the \gls{Fourier transform} is used. The algorithm used to compute the discrete \gls{Fourier transform} is called \gls{FFT}. The \gls{frequency spectrum} calculated by \gls{FFT} is discrete---if the discrete real-valued input signal, as is the case with \gls{EEG} recording, has $N$ values then the output has $\frac{N}{2}$ values. This derives from the definition and symmetric property of the discrete \gls{Fourier transform}. %The higher the sampling rate and \gls{ADC} resolution of a recording device, the more accurate the recording and therefore the more accurate the frequency spectrum and the periodogram of the signal.

The signal that is recorded, however, may contain frequencies that are too high to be detected from the discrete signal. Theoretically, the \gls{sampling rate} of the device has to be more than two times higher than the highest frequency in the signal to reconstruct the continuous signal from the discrete signal and also to decompose the signal into \glspl{frequency component}. When looking it from different angle, the highest frequency that can be detected with a \gls{sampling rate} of $f$ is $\frac{2}{f}$ and it is called \gls{Nyguist frequency}. But since real-world application are imperfect, even higher \gls{sampling rate} is needed.

To conclude previous discussion, the \gls{frequency spectrum} calculated by \gls{FFT} from real-valued time-domain signal with $N$ values is defined at frequencies $\frac{1}{N}, \frac{2}{N}, \dots\frac{f}{2N}$. These frequencies are also called \glspl{frequency bin}. The length of a \gls{frequency spectrum} depends on the length of the signal from which it is calculated. The longer the input signal, the more \glspl{frequency bin} will be acquired.

Thus a time-domain signal can be decomposed into \glspl{pure tone} or sine and cosine waves using \gls{FFT}. \gls{FFT} calculates the \gls{frequency spectrum} of a time-domain signal. \Gls{frequency spectrum} can be converted to the estimation of \gls{power spectral density} which contains only the information about the amplitude of the \glspl{pure tone}. This knowledge can be used to detect \glspl{SSVEP} in \gls{EEG} recording.

%As already discussed in section~\ref{sec:EEG_comparison}, Emotiv EPOC has high internal sampling rate to filter out high frequencies

%Thus higher sampling rate and \gls{ADC} resolution lead to more accurate \gls{SSVEP} detection.

\section{Decomposition of target flickering}

This section will focus on the decomposition of the \glspl{flickering waveform}. The \gls{target} \gls{flickering} and decomposition process were discussed in section~\ref{sec:stimuli} and section~\ref{sec:fourier} respectively. As discussed in previous section, only sine and cosine waves are composed of single frequency. Other waveforms have more \gls{frequency component}. The \gls{frequency component} with lowest frequency is called \gls{fundamental} or the first \gls{harmonic} of the signal. \Gls{harmonic} of a signal is a \gls{frequency component} with frequency that is integer multiple of the \gls{fundamental} frequency of the signal. If the \gls{fundamental} frequency is $f$, then first, second are third \glspl{harmonic} have frequencies of $1f$, $2f$, $3f$ respectively.

It can be shown that a \gls{square wave} is composed of odd \glspl{harmonic}. If the \gls{flickering waveform} is \gls{square wave} or \gls{rectangular wave}, the \gls{fundamental} frequency of the waveform is the frequency of the stimuli presentation. Therefore, \gls{square wave} with \gls{fundamental} frequency $f$ can be represented as a sum of sine waves
\begin{equation}
	\label{eq:square}
	square(t) = \sum_{n=1,3,5,\dots}^{\infty}\frac{1}{n} \sin(2\pi nft)
\end{equation}
where $square(t)$ is the \gls{flickering} represented as \gls{state} versus time, $\frac{1}{n}$ is the amplitude of a \gls{frequency component} and $f$ is the \gls{fundamental} frequency of the waveform. It can be seen that the \gls{fundamental} has the highest amplitude. \Gls{rectangular wave}'s \glspl{frequency component} depend on the \gls{duty cycle} of the wave. But as is the case with \glspl{square wave}, the \gls{fundamental} has the highest amplitude also in \gls{rectangular wave}.

Unfortunately, the \gls{SSVEP} response to \gls{target} \gls{flickering} does not contain only the frequencies that are present in the \glspl{frequency component} of the \gls{flickering waveform}---\gls{SSVEP} contains other frequencies too. However, the frequencies that are present in the \glspl{frequency component} of the \gls{flickering waveform} are more successfully elicited in \gls{SSVEP} response~\cite{square_sine}. It has even been reported that \gls{SSVEP} contains small \glspl{subharmonic} of the \gls{flickering waveform}~\cite{ssvep_response}. \Gls{subharmonic} of a signal is \gls{frequency component} with frequency $\frac{f}{n}$, where $f$ is the \gls{fundamental} frequency of the signal and $n$ is an integer.

\gls{SSVEP} contains \glspl{harmonic} of the \gls{flickering} frequency~\cite{ssvep_response}. Therefore, \gls{target} frequencies should be chosen so that the frequencies are not each other \glspl{harmonic} or \glspl{subharmonic}. Otherwise it might not be distinguishable which \gls{target} \gls{flickering} elicits which frequency in the \gls{SSVEP} response since. For example, if 6 Hz ans 12 Hz \gls{flickering} frequencies are used, then 6 Hz \gls{flickering} also elicits 12 Hz frequency in the \gls{SSVEP} since 12 Hz is the second harmonic of 6 Hz \gls{flickering}.

Thus \gls{SSVEP} reflects certain properties of the visual stimulus. The frequencies that are present in \gls{flickering waveform} are likely to be found also in the \gls{SSVEP} response, but other \glspl{frequency component} are present in the \gls{SSVEP} too~\cite{square_sine}. It is sufficient to detect only the \gls{frequency component} with the same frequency as the \gls{target} \gls{flickering}, but to improve the performance other \glspl{frequency component} should be detected too~\cite{harmonic_imrpovement}.

%. Most important of these properties is that \gls{SSVEP} has a component with the same frequency as the visual stimulus. But that is not the only component frequency in \gls{SSVEP}. \gls{SSVEP} has other components too that are discussed in section~\ref{sec:fourier}. 

\section{Related work}

There are many types of \glspl{BCI} available. A review by Bashashati \textit{et al.}~\cite{bci_comparison} contains a detailed overview of \gls{EEG}-based \glspl{BCI}. Their paper includes overview of the following neuromechanisms used in \gls{EEG}-based \glspl{BCI}: \gls{VEP} and \gls{SSVEP}, P300, slow cortical potential, response to mental task, sensorimotor activity, and multiple neuromechanisms or hybrid \gls{BCI}. Hybrid \gls{BCI} uses two different neuromechanisms in a \gls{BCI} and therefore two different methods are required to analyse \gls{EEG} recording~\cite{hybrid_bci, hybrid_bci2}. %In this thesis, main focus is on \gls{SSVEP}-based \glspl{BCI}. 

\gls{SSVEP}-based \glspl{BCI} can be divided in categories according to the method used to detect \glspl{SSVEP} in \gls{EEG} recording. These methods are also called \gls{feature extraction} methods. Current \gls{SSVEP}-based \gls{BCI} \gls{feature extraction} methods include:
\begin{itemize}
	\item \Gls{PSDA} method introduced by Cheng \textit{et al.}~\cite{psda}.
	\item \Gls{SC} method introduced by Wu and Yao~\cite{sc}.
	\item Dual-frequency \gls{SSVEP} methods~\cite{dual1, dual2}.
	\item \Gls{MPCC} method introduced by Tong \textit{et al.}~\cite{MPCC}.
	\item \Gls{MEC} method introduced by Friman \textit{et al.}~\cite{mec}. This method has shown better performance than \gls{SC} and \gls{PSDA} method~\cite{mec_comparison}.
	\item \Gls{CCA} method introduced by Lin \textit{et al.}~\cite{cca_lin}. This method has shown better performance than \gls{PSDA} method~\cite{cca_psda, bin2009cca, cca_lin}.
	\item \Gls{LASSO} method introduced by Zhang \textit{et al.}~\cite{LASSO}. This method has shown better performance than \gls{CCA} method~\cite{LASSO}.
	\item \Gls{LRT} method introduced by Zhang \textit{et al.}~\cite{LRT}. This method has shown better performance than \gls{CCA} method and similar performance to \gls{LASSO} method~\cite{LRT}.
\end{itemize}
There is also an improvement of \gls{CCA} method available called multiway \gls{CCA}, which has shown slightly better performance than standard \gls{CCA} method~\cite{mcca}. For more comprehensive review see article by Liu \textit{et al.}~\cite{feature_extraction}.

In this thesis, Emotiv EPOC is used to record brain activity. The following papers also describe \glspl{BCI} that used Emotiv EPOC to record brain activity:
\begin{itemize}
	\item Liu \textit{et al.}~\cite{emotiv_11hz} used \gls{CCA} method;
	\item Lin \textit{et al.}~\cite{emotiv_walking} used \gls{CCA} method;
	\item Choi and Jo~\cite{emotiv_hybrid} designed hybrid \gls{BCI};
	\item Zier~\cite{emotiv_psda} used \gls{PSDA} method;
	\item Hvaring and Ulltveit-Moe~\cite{emotiv_comparison} compared different \gls{feature extraction} methods;
	\item Duvinage \textit{et al.}~\cite{emotiv_p300_comp} used P300 method and compared the performance of Emotiv EPOC and a medical-grade device.
\end{itemize}
%This thesis focuses on two of these methods: \gls{PSDA} and \gls{CCA} methods.


\section{SSVEP-based brain-computer interface}
\label{sec:SSVEP_detection}
The aim of this chapter is to describe two methods used to detect \glspl{SSVEP} in \gls{EEG} recording. These methods are called \gls{feature extraction} methods. This section describes the \gls{PSDA} and \gls{CCA} \gls{feature extraction} methods. These are the methods used in chapter~\ref{sec:SSVEP_BCI} to design application for controlling a robot. 

\subsection{Power spectral density analysis}

\Gls{PSDA} is widely used in \gls{SSVEP}-based \glspl{BCI}~\cite{bin2009cca}. This method is based on a \gls{power spectral density} estimation. \Gls{power spectral density} estimation method called \gls{periodogram} was discussed in section~\ref{sec:fourier}.

There are different ways how to use the \gls{power spectral density} estimation for \gls{feature extraction}. Some \glspl{BCI} use a training session to find a threshold value that certain frequency's amplitude or power has to exceed in order to select the \gls{target} with corresponding frequency as user's choice. This method requires a training session during which the threshold values are determined, but \gls{SSVEP}-based \glspl{BCI} could be implemented without the need for a training session. Threshold values, however, can be used together with other methods that \gls{power spectral density}. For example, article by Lin \textit{et al.} describes \gls{BCI} that uses peak finding to find highest values in \gls{power spectral density}~\cite{cca_lin}.

Common method in \glspl{BCI} that use \gls{PSDA} \gls{feature extraction} is to calculate \gls{SNR} or other values that are related to \gls{SNR}. For example, the ratio of frequency's power to the mean of adjacent frequency bins can be calculated
\begin{equation}
	\label{eq:snr1}
	\mbox{SNR}(f_k) = \frac{2P(f_k)}{P(f_{k+1})+P(f_{k-1})}
\end{equation}
where $\mbox{SNR}(f_k)$ is the \gls{SNR} of frequency $f_k$, $k$ is the number of the \gls{frequency bin} in \gls{periodogram}, assuming that the \glspl{frequency bin} are numbered in order of increasing frequencies and $P$ is the function representing \gls{power spectral density}~\cite{psda_snr1}. The \gls{SNR} as defined in equation~\ref{eq:snr1} can be calculated for every \gls{target} frequency and the \gls{target} that has the frequency with highest \gls{SNR} is assumed to be the user's choice. More than two adjacent frequency bins can be used to calculate the \gls{SNR}. Therefore, the equation~\ref{eq:snr1} can be generalised
\begin{equation}
	\mbox{SNR}(f_k) = \frac{nP(f_k)}{\sum_{i=1}^{n/2}P(f_{k-i})+P(f_{k+i})}
\end{equation}
where n is the number of adjacent frequency bins used~\cite{psda_snr2}. To generalise it even more, the ratio of the frequency's power to the whole \gls{power spectral density} or all the \glspl{frequency bin} can be calculated
\begin{equation}
	\mbox{SNR}(f_k) = \frac{P(f_k)}{\sum_i P(f_i)}
\end{equation}
This is the same as using normalised \gls{power spectral density} and comparing the normalised powers. It is also possible to just compare the powers of different \gls{target} frequencies and select the \gls{target} with highest power~\cite{cca_psda}. All these methods can use in addition to the \gls{target} frequency also its \glspl{harmonic}. In this case the sum of the \glspl{SNR} can be used
\begin{equation}
	\mbox{SNR}(f_k, 2f_k, \dots, hf_k)=\sum_{i=1}^{h}\mbox{SNR}(if_k)
\end{equation}
where $h$ is the number of \glspl{harmonic} used. Often three \glspl{harmonic} are used. After calculating the \gls{SNR} for all the \gls{target} frequencies and for their \glspl{harmonic} if necessary, the \gls{target} with highest \gls{SNR} or the highest sum of \glspl{SNR} can be selected as user's choice.

One problem that occurs is that the \gls{power spectral density} is discrete. \Gls{power spectral density} might not have frequency bins on the exact frequencies of interest---the frequencies of \glspl{target} used in \gls{BCI}. In this case \gls{interpolation} can be used~\cite{cca_psda}. \Gls{interpolation} is used to construct a discrete signal between the discrete points or in other words to approximate the continuous signal from which the discrete values were extracted from. Another possibility to solve this problem is to use \gls{zero padding}. \Gls{zero padding} means that zeros are added to the end of the discrete signal before performing \gls{FFT}. This results in more \glspl{frequency bin} in the \gls{power spectral density}, because the number of \glspl{frequency bin} depends on the length of the input signal. The only alteration in \gls{power spectral density} is that it has more \glspl{frequency bin} if acquired from zero-padded signal.

To design a \gls{SSVEP}-based \gls{BCI} that uses \gls{PSDA} \gls{feature extraction} method it is enough to use one of the methods described in this section. Often \glspl{SNR} or powers of different \glspl{target} are compared to determine the user's choice. Some signal processing techniques that are required or can be used to enhance the performance of a \gls{BCI} will be discussed in section~\ref{sec:signal_pipeline}.

\subsection{Canonical correlation analysis}

\Gls{CCA} was first introduced by Harold Hotelling in 1936~\cite{cca_hotelling}. In 2001 \gls{CCA} was used to introduce a novel method for detecting neural activity in \gls{fMRI} data~\cite{cca_fmri}. Likewise, \gls{CCA} was introduced to \gls{EEG} recording analysis for the first time in 2007~\cite{cca_lin}. 

\Gls{CCA} is statistical method that can be used when analysing the correlation between two sets of data. In \gls{SSVEP}-based \glspl{BCI} one set of data is the multichannel \gls{EEG} recording. For example, if recording data with electrodes located in O1 and O2, the recorded data can be represented with canonical variable~$X=x_{O1}+x_{O2}$. The second set of data is the \gls{fundamental} frequency and \glspl{harmonic} of the \gls{flickering} frequency. Both sine and cosine waves are used in the second data set % since the phase of the \gls{SSVEP} is not known.
\begin{equation}
	\label{eq:cca_ref}
	Y=\left(\begin{matrix}
		sin(2\pi)\\
		cos(2\pi)\\
		sin(2\pi)\\
		cos(2\pi)\\
		sin(2\pi)\\
		cos(2\pi)\\
	\end{matrix}\right)
\end{equation}
In the method proposed by ... \textit{et al.} three harmonics are used as in equation~\ref{eq:cca_ref}.
